Manuscript_v0.5_genetic_covariances

### 2020-11-13 R/Manuscript_v0.5_genetic_covariances.Rmd
Bivariate genetic covariances in Gulf subpop

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(asreml)
library(cowplot)
library(switchgrassGWAS)
```

# Load data & prep multivariate model data frames.
Add any data that needs to be loaded for these analyses here.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

@@

@@

@@

8+7+6+5+4+3+2+1 # 36 var and covar in eight garden model.
# programmatically, each pair of two gardens, k and k+1. (special case of 8 & 1 also).
```
check model ran and saved ok on the server
```{r}
asr_8 <- readRDS("../analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
summary(asr_8)$varcomp
vpredict(asr_8, h2_tx1 ~ V1/(V1+V38))
vpredict(asr_8, h2_tx2 ~ V3/(V3+V40))
vpredict(asr_8, h2_tx3 ~ V6/(V6+V43))
vpredict(asr_8, h2_ok ~ V10/(V10+V47))  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
vpredict(asr_8, h2_mo ~ V15/(V15+V52))
vpredict(asr_8, h2_ne ~ V21/(V21+V58))
vpredict(asr_8, h2_mi ~ V28/(V28+V65))
vpredict(asr_8, h2_sd ~ V36/(V36+V73))

vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5))  # sig pos

vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5))   # sig pos
vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5))   # sig pos

vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) 
vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5)) 
vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))

vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) 
vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5))  # ** sig neg even plus 2SE
vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) 
vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5))

vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) 
vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) 
vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) 
vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) 
vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5))   # sig pos

vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) 
vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) 
vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) 
vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) # sig pos
vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5))

vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) 
vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) # sig pos
vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) 
vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5))

#36.80566-(16.49249*2)
```
36.80566-(16.49249*2)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(asreml)
library(cowplot)
library(switchgrassGWAS)
```

# Load data & prep multivariate model data frames.
Add any data that needs to be loaded for these analyses here.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
inputdir <- file.path(workingdir, "analysis", "genetic-covariances")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

@@

# Will Midwest models converge for bivariate analyses?

Can't get trivariate analysis to converge ever for the Midwest subpop, though the Gulf & G&M models do converge. 
Unfortunately I will need to fit ~28 models per phenotype and subpop for bivariate models to get all of the genetic covariances. Blehhhhh.
```{r}
garden_v <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD")
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
                 Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
k=3; j = 1; i = 1; q = 1
var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop", Garden_1 = "garden1", Garden_2 = "garden2", loglik = 0)

append_vpredict_bivar <- function(vpredict_df, var_df, 
                                  phenotype = names(phe_df)[j], 
                                  subpop = names(subpop_v)[i], 
                                  garden1 = garden_v[k],
                                  garden2 = garden_v2[q], loglik = mloglik){
  var_df <- vpredict_df %>%
    as_tibble(rownames = "Value") %>%
    mutate(Phenotype = phenotype,
           Subpop = subpop,
           Garden_1 = garden1,
           Garden_2 = garden2,
           loglik = mloglik) %>%
       add_row(var_df, .)
  return(var_df)
}

for(i in seq_along(subpop_v)){
  for(j in seq_along(phe_df)){
    for(k in seq_along(garden_v)){
      garden_v2 <- garden_v[-(1:k)]
      for(q in seq_along(garden_v2)){
    cov_df_use <- phe_df[[j]] %>%
  filter(SUBPOP %in% subpop_v[[i]]) %>%
      rename(garden1 = garden_v[k], garden2 = garden_v2[q]) %>%
      filter(!(is.na(garden1) & is.na(garden2)))
    tryCatch({
      asr_2 <- asreml(cbind(garden1, garden2) ~ trait,
                      random = ~us(trait, init = c(1, 0.1, 1)):vm(PLANT_ID,
                                                                    k_full), 
                      residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                      na.action = na.method(x = "omit"),
                      data = cov_df_use, 
                      workspace = "8gb", maxiter = 300)
      asr_2 <- update.asreml(asr_2, maxiter = 20)
      saveRDS(asr_2, file.path(inputdir, paste0(names(phe_df)[j], "_asreml_cov_model_", garden_v[k], "_", garden_v2[q], "_gardens_", names(subpop_v)[i], "_subpop_200_iterations.rds")))
      mloglik = asr_2$loglik[1]
      var_table <- vpredict(asr_2, h2_garden1 ~ V1/(V1+V5)) %>% 
        append_vpredict_bivar(., var_table)
      var_table <- vpredict(asr_2, h2_garden2 ~ V3/(V3+V7)) %>% 
        append_vpredict_bivar(.,var_table)
      var_table <- vpredict(asr_2, r_G ~ V2/((V1*V3)^0.5)) %>% 
        append_vpredict_bivar(., var_table)
      }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
    saveRDS(var_table, file.path(inputdir,     
                                 "Two_site_var_table_up_to_300_iterations.rds"))
      }
    }
  }
}

```

```{r}
var_table <- read_rds(file.path(inputdir,     
                                "Two_site_var_table_up_to_300_iterations.rds"))

var_table %>%
  write_csv(., file.path(inputdir,     
                                "Two_site_var_table_up_to_300_iterations.csv"))
```

@@

#------------
# Three site covariances loop
```{r}
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
                 Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)

var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop")

for(i in seq_along(subpop_v)){
  for(j in seq_along(phe_df)){
    cov_df_use <- phe_df[[j]] %>%
  filter(SUBPOP %in% subpop_v[[i]])
      asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
                      random = ~us(trait, init = c(1, 0.1, 1, 
                                                   0.1, 0.1, 1)):vm(PLANT_ID,
                                                                    k_full), 
                      residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 
                                                           0.1, 0.1, 0.1)), 
                      na.action = na.method(x = "omit"),
                      data = cov_df_use, 
                      workspace = "6gb", maxiter = 2600)
      saveRDS(asr_3, paste0("analysis/genetic-covariances/", names(phe_df)[j], "_asreml_cov_model_3_gardens_", names(subpop_v)[i], "_subpop_2600_iterations.rds"))

     var_table <- vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .)
     var_table <- vpredict(asr_3, h2_MI ~ V3/(V3+V10)) %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .)
     var_table <- vpredict(asr_3, h2_MO ~ V6/(V6+V13)) %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .) 
     var_table <- vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .) 
     var_table <- vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5))  %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .)
     var_table <- vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) %>% 
       as_tibble(rownames = "Value") %>%
       mutate(Phenotype = names(phe_df)[j],
              Subpop = names(subpop_v)[i]) %>%
       add_row(var_table, .) 
      saveRDS(var_table, "Three_site_var_table_50_iterations_usually.rds")

  }
}
asr_3 <- update.asreml(asr_3, maxiter = 100)
```

## view ASReml output from three site covariance models
```{r}
var_table <- read_rds(file.path(inputdir,
                          "Three_site_var_table_50_iterations_usually_v2.rds"))
var_table %>% write_csv(., file.path(inputdir, "Three_site_var_table_50_iterations_or_more.csv"))
```

@@

8+7+6+5+4+3+2+1 # 36 var and covar in eight garden model.
# programmatically, each pair of two gardens, k and k+1. (special case of 8 & 1 also).
```
check models ran and saved ok on the server

## View 8 output and save to a var_table
```{r}
asr_8 <- readRDS("../analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
asr_8 <- readRDS("../analysis/genetic-covariances/FL50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
summary(asr_8)$varcomp

subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
                 Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
j=1
i=1
summary(asr_8)$call
var_table <- tibble(Value = "Value", Estimate = 0, SE = 0, Phenotype = "Phenotype", Subpop = "Subpop")
append_vpredict <- function(vpredict_df, var_df, phenotype = names(phe_df)[j], 
                            subpop = names(subpop_v)[i]){
  var_df <- vpredict_df %>%
    as_tibble(rownames = "Value") %>%
    mutate(Phenotype = phenotype,
           Subpop = subpop) %>%
       add_row(var_df, .)
  return(var_df)
}

var_table <- vpredict(asr_8, h2_tx1 ~ V1/(V1+V38)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_tx2 ~ V3/(V3+V40)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_tx3 ~ V6/(V6+V43)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_ok ~ V10/(V10+V47)) %>% 
  append_vpredict(., var_table)  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
var_table <- vpredict(asr_8, h2_mo ~ V15/(V15+V52)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_ne ~ V21/(V21+V58)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_mi ~ V28/(V28+V65)) %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, h2_sd ~ V36/(V36+V73)) %>% 
  append_vpredict(., var_table)

var_table <- vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5)) %>% 
  append_vpredict(., var_table)  # sig pos

var_table <- vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5)) %>% 
  append_vpredict(., var_table)   # sig pos
var_table <- vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5)) %>% 
  append_vpredict(., var_table)   # sig pos

var_table <- vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5))  %>% 
  append_vpredict(., var_table)
var_table <- vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))  %>% 
  append_vpredict(., var_table)

var_table <- vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5)) %>% 
  append_vpredict(., var_table)  # ** sig neg even plus 2SE
var_table <- vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5)) %>% 
  append_vpredict(., var_table)

var_table <- vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5)) %>% 
  append_vpredict(., var_table)   # sig pos

var_table <- vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) %>% 
  append_vpredict(., var_table) # ** sig neg even plus 2SE
var_table <- vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) %>% 
  append_vpredict(., var_table) # sig pos
var_table <- vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5)) %>% 
  append_vpredict(., var_table)

var_table <- vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) %>% 
  append_vpredict(., var_table) # ** sig neg even plus 2SE
var_table <- vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) %>% 
  append_vpredict(., var_table) # ** sig neg even plus 2SE
var_table <- vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) %>% 
  append_vpredict(., var_table) # ** sig neg even plus 2SE
var_table <- vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) %>% 
  append_vpredict(., var_table) # sig pos
var_table <- vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) %>% 
  append_vpredict(., var_table) 
var_table <- vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5)) %>% 
  append_vpredict(., var_table)

write_csv(var_table, file.path(inputdir, "Var_Table_8_garden_Gulf_models_20_iter.csv"))
#36.80566-(16.49249*2)
```
36.80566-(16.49249*2)
